---
title: "Credit Card Customer Churn"
author: "Adogbeji Agberien"
date: "13/05/2021"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = F, warning=F}
# Clear working directory
rm(list = ls())

# Import and load packages
required_packages <- c("RColorBrewer", "cowplot", 
                       "lubridate", 
                       "Hmisc", "psych",
                       "tidyverse", "data.table", 
                       "mlr3")

packageCheck <- lapply(required_packages, FUN = function(x) {
  if(!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})
```

# Background 

Customer churn is the loss/turnover of a client. For this analysis, the goal is to predict/classify customers who will churn.

```{r}
# Import data 
bank_churn <- fread("C:/Users/diji_/Desktop/Data Science/Projects/Bank Churn/BankChurners.csv")
```

```{r include = F}
bank_churn <- bank_churn[, c(-22, -23)]
```

```{r}
# Show the first 3 rows of the data 
head(bank_churn, 3)
```
We have some bit of features here, the first task will be to identify what the output feature is. In this case it is the Attrition_Flag. As shown below, the customers are either existing or "Attrited"; probably meaning they are no longer customers/clients. 

```{r}
unique(bank_churn$Attrition_Flag)
```

Next step is some Exploratory Data Analysis (EDA) and pre-processing. Thus far, the only input feature I see that is not necessary to be included in our model is the "CLIENTNUM", The CLIENTNUM is just an identifier. I won't remove it just yet though, in accordance with the theme of EDA I'll make a plot to see the relationship between the client ID and Attrition Flag. We have no time variable to evaluate the churning through time. So, my maybe flawed logic (*No bank manager to talk to lol*) is that the recent customers have higher client numbers than earlier customers...you know, customer 1, customer 2, ...., customer *n*, etc. 

Let us encode the response variable with 1 representing that the customer churned and 0 representing otherwise. 

```{r}
# Encode the response 
bank_churn$Attrition_Flag <- factor(bank_churn$Attrition_Flag, levels = c("Existing Customer", "Attrited Customer"), 
                                    labels = c(0, 1))
```

```{r}
ggplot() +
  geom_point(aes(x = bank_churn$CLIENTNUM, y = bank_churn$Attrition_Flag)) +
  xlab("Client Number") +
  ylab("Churn (1) or No churn (0)") +
  theme_classic()
```
Damn, no cigar! 

Let us find out what the other features mean so we can make sense of our data as we work through the challenge. 

Card_Category: 

Months_on_book: 

Total_Relationship_Count: 

Months_Inactive_12_mon: 

Contacts_Count_12_mon: 

Total_Revolving_Bal: 

Avg_Open_To_Buy: 

Total_Amt_Chng_Q4_Q1:  

Total_Trans_Amt:  

Total_Trans_Ct:  

Total_Ct_Chng_Q4_Q1:  

Avg_Utilization_Ratio: 